The Weakly Nonlinear Saturation of the Magnetorotational Instability

### 2016-07-15 paper/mri_paper.tex
edited norm of allorders to not include magic number. Still includes real imag shift
fixed N2 psi terms in paper
fixed N2 u terms in paper
Pm 2E-3
removed hmri section from thin gap paper

\title{The Weakly Nonlinear Saturation of the Magnetorotational Instability}
\author{Clark, S.E.\altaffilmark{1}}
\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
\altaffiltext{3}{Department of Physics, SUNY Farmingdale}

@@

% Jeff's sims!!!!!

\section{Helical MRI}

The configuration investigated thus far, that of magnetized Taylor Couette flow with a purely axial magnetic field, constitutes the standard MRI. The helical MRI (HMRI) is a related, but separate, MHD instability, first identified by Hollerbach and R\"udiger (\citeyear{Hollerbach:2005tr}). It arises in the presence of combined axial and azimuthal fields.

We initialize the helical MRI by adding a constant background toroidal field to Equation \ref{backgroundfield1}, such that our initial magnetic field structure becomes

\beq
\mathbf{B} = B_0 \hat{z} + \xi B_0 \hat{y},
\eeq

where $\xi$ is a dimensionless parameter representing the relative strength of the axial field and the vertical field.

This addition leaves Equations \ref{eqset2} - \ref{eqset4} unchanged, but Equation \ref{eqset1} gains a term to become

\begin{multline}
\label{eqset1_hmri}
\partial_t \nabla^2 \Psi \, = \, \frac{2}{\beta} B_0 \partial_z \nabla^2 A \, + \, 2 \partial_z u_{y} \, - \xi B_0 \partial_x \partial_z B_{y}\\
- \, J\left(\Psi, \nabla^2 \Psi\right) \, + \, \frac{2}{\beta}J\left(A, \nabla^2 A \right) \, + \, \frac{1}{\reye}\nabla^4 \Psi,
\end{multline}

in other words, the $\Psi$ equation picks up a linear $B_y$ term.

Following the same prescription for weakly nonlinear tuning, multiscale expansion, and perturbation expansion, we find the equations for the HMRI up to order $\epsilon^3$.

To $\mathcal{O}(\epsilon)$, we find

\beq
\mathcal{L} \mathbf{V_1} + \xi B_0 \mathcal{H} \partial_z \mathbf{V_1} = 0
\eeq

\beq
\mathcal{L}\mathbf{V_2} = \mathbf{N_2} - \widetilde{\mathcal{L}}_1 \partial_Z \mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_2} - \xi \mathcal{H}\partial_Z \mathbf{V_1}
\eeq

\begin{multline}
\mathcal{D}\partial_T \mathbf{V_1} \, + \, \mathbf{N_3} \, = \, \mathcal{L} \mathbf{V_3} \, + \, \widetilde{\mathcal{L}}_1\partial_Z\mathbf{V_2} \, + \, \widetilde{\mathcal{L}}_2\partial_Z^2\mathbf{V_1} \, \\
- \partial_z\mathcal{X}\mathbf{V_1} \, - \partial_z^3\mathcal{L}_3\mathbf{V_1} + \xi \mathcal{H} \partial_z \mathbf{V_3} - \xi \mathcal{H} \partial_Z \mathbf{V_2} + \xi \mathcal{H} \partial_z \mathbf{V_1}.
\end{multline}

In the above,

\beq
\mathcal{H} = \left[\begin{matrix}
0 & 0 & 0 & \partial_x \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\
0 & 0 & 0 & 0 \\ \end{matrix} \right].
\eeq

The first amplitude equation (\ref{ampeq1}) therefore becomes

\begin{multline}
\label{ampeq1_hmri}
a \partial_T \alpha \, = - \, c \alpha \left|\alpha^2\right| \, - \, \widetilde{c} \alpha \partial_Z \beta \,  - b i k_c \alpha \, + \, h \partial_Z^2 \alpha \, \\
+ \, g i k_c^3 \alpha \, + \, q \xi \partial_Z^2 \alpha \, + \, w i k_c \xi \alpha,
\end{multline}

\beq
q = \left< \mathbb{V}^\dagger \cdot \left(\mathcal{H} \mathbb{V}_{21}\right)^* \right>,
\eeq

\beq
w = \left< \mathbb{V}^\dagger \cdot \left(\mathcal{H} \mathbb{V}_{11}\right)^*\right>.
\eeq

The second amplitude equation (\ref{ampeq2}) becomes

\beq\label{ampeq2_hmri}
\partial_T \beta = - m \partial_Z^2 \beta + \frac{1}{\reym} \partial_Z^2 \beta \, - \, s \xi 2 \left|\alpha\right|\partial_Z \alpha \, - \, f\xi\partial_Z^2 \beta,
\eeq

\beq
s = \left<\mathbb{U}_{11} \cdot \left(\mathcal{H}\mathbb{V}_{20}\right)^*\right>
\eeq

\beq
f = \left< \mathbb{U}_{11} \cdot \left(\mathcal{H}\mathbb{U}_{20}\right)^*\right>.
\eeq

If $s = 0$, as it is for the standard MRI, the two amplitude equations will again decouple.

@@

\beq
\begin{split}
N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c \partial_x^2 \Psi_{11}^* - i k_c^3 \Psi_{11}^*\right) \\
& \frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c \partial_x^2 A_{11}^* - i k_c^3 A_{11}^*\right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
\end{split}
\eeq

\beq
N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* - \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* + \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
\eeq

@@

\beq
N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* - \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* + \partial_x A_{11} \cdot i k_c u_{11}^*
\eeq \\

@@

\beq
\begin{split}
N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
 + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
 & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
\end{split}
\eeq

>>>>

\title{The Weakly Nonlinear Saturation of the Magnetorotational Instability}
\author{Clark, S.E.\altaffilmark{1}}
\author{Oishi, J.S. \altaffilmark{2,}\altaffilmark{3}}
%\author{Mac Low, M.-M.\altaffilmark{1,}\altaffilmark{2}}
\altaffiltext{1}{Department of Astronomy, Columbia University, New York, NY} 
\altaffiltext{2}{Department of Astrophysics, American Museum of Natural History, New York, NY}
\altaffiltext{3}{Department of Physics, SUNY Farmingdale}

@@

@@

\beq
\begin{split}
N_{20}^{(\Psi)} = & \, i k_c \Psi_{11} \cdot \left(\partial_x^3 \Psi_{11}^* - k_c^2 \partial_x \Psi_{11}^*\right) - \partial_x \Psi_{11} \cdot \left(i k_c^3 \Psi_{11}^* - i k_c \partial_x^2 \Psi_{11}^*\right) \\
& +\frac{2}{\beta} \partial_x A_{11} \cdot \left(i k_c^3 A_{11}^* - i k_c \partial_x^2 A_{11}^* \right) - \frac{2}{\beta} i k_c A_{11} \cdot \left(\partial_x^3 A_{11}^* - k_c^2 \partial_x A_{11}^*\right)
\end{split}
\eeq

\beq
N_{20}^{(u)} = i k_c \Psi_{11} \cdot \partial_x u_{11}^* + \partial_x \Psi_{11} \cdot i k_c u_{11}^* - \frac{2}{\beta} i k_c A_{11} \cdot \partial_x B_{11}^* - \frac{2}{\beta} \partial_x A_{11} \cdot i k_c B_{11}^*
\eeq

@@

\beq
N_{20}^{(B)} = i k_c \Psi_{11} \cdot \partial_x B_{11}^* + \partial_x \Psi_{11} \cdot i k_c B_{11}^* - i k_c A_{11} \cdot \partial_x u_{11}^* - \partial_x A_{11} \cdot i k_c u_{11}^*
\eeq \\

@@

%Correct -- checked 11/14/15
\beq
\begin{split}
N_{31}^{(A)} =\, & -i k_c \left(A_{11}\cdot\partial_x \Psi_{20}\right) - i k_c \left(A_{11} \cdot \partial_x \Psi_{20}^*\right)
 + i k_c \left(A_{11}^* \cdot \partial_x \Psi_{22}\right) + i 2 k_c \left(\partial_x A_{11}^* \cdot \Psi_{22}\right) \\
 & + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}\right) + i k_c \left(\Psi_{11} \cdot \partial_x A_{20}^*\right) - i k_c \left(\Psi_{11}^* \cdot \partial_x A_{22} \right) - i 2 k_c \left(\partial_x \Psi_{11}^* \cdot A_{22}\right)
\end{split}
\eeq
